Seminarium TI 3700\\ Data Analysis on Twitter: An Overview (working title)

### 2014-03-11 authorinstr_2.tex
Intro

@@

\item \textbf{Pair-Wise Interaction Tensor Factorization} The Pair-Wise Interaction Tensor Factorisation is a state-of-the-art algorithm that has
been developed for the purpose of tag-based recommendation systems. The model can be adapted to compute recommendations from tags used by Twitter.

>>>>

Microblogging has grown out to be one of the most interesting forms of social media. One of its most popular implementations has to be Twitter, where
users can share short messages, called Tweets, with every other user. The limitation in characters forces users to 
compress all their information into a small message, making them extra interesting to analyse. These Tweets often contain information broadcasted by the user in
the form of opinions, emotions, interests, etc. One of the biggest challenges is finding a way to extract the intended information from these Tweets. Once
the information is extracted however, the possible applications are endless. This paper is dedicated to combining succesful techniques and useful applications
of these social web data analytics. In the following section, Twitters API gets introduced which will show that the platform already offers great tools
and commands to retrieve specific datasets, or a completely random sample stream of Tweets. After that, two entire sections will be dedicated to all sorts
of techniques that are being used to extract this information. One section will focus on single tweets, and the following on sets of multiple Tweets. Both
of these sections will involve several usecases for certain techniques.

%TODO: Add what the conclusion will say

@@

\item \textbf{Pair-Wise Interaction Tensor Factorization}

The Pair-Wise Interaction Tensor Factorisation is a state-of-the-art algorithm that has
been developed for the purpose of tag-based recommendation systems. The model can be adapted to compute recommendations from tags used by Twitter. 
The adaptation involves changing the roles of resources and tags with respect to each other.

Our adaptation of the model to resource
recommendation simply exchanges the roles of resources and tags with respect to each
other. This model-based approach generates a set of factor matrices which resembles
a special case of the Tucker decomposition of a tensor. The tensor itself is not directly
induced by the data, but rather reflects a ranking over the resources for each user-tag
pair. Thus, it is important to note that this model is applicable only to the special case of
tag-based resource recommendation, and not for resource recommendation tasks with
other requirements.
The model is built by considering observations in the data of the form (u, r+ , r− , t),
where (u, r+ , t) is a triple which is found in the data (a positive example of resource
selection) and (u, r− , t) is a triple not found in the data (a negative example of re-
source selection). An iterative gradient-descent algorithm is employed to optimize a116
J. Gemmell et al.
ranking function that prefers positive examples in the data over negative ones. Each of
four related matrices is updated until convergence is found. The matrices represent the
U
T
factor-reduced components of the specialized tensor factorization M = Uk Rk
+Tk Rk
,
U
where Uk is the user factor matrix, Tk is the tag factor matrix, Rk is the resource factor
T
matrix with respect to users and Rk
is the resource factor matrix with respect to tags, k
is the selected number of factors, and M is the personalized resource-ranking tensor.
Generating a resource recommendation for a given user u and tag t is simply a
matter of referring to the appropriate user-tag column of the ranking tensor M . The
relevance score of a resource given a user-tag pair is calculated as: φ(u, {t}, r) =
